\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control}

### 2018-01-21 paper/final/Response_to_reviewers/Response_to_reviewers.tex

\section{Response to Associate Editor}

\begin{enumerate}

	\item \textbf{Associate Editor:} \textit{The authors should address the points and clarifications in the reviews, and also specifically address the novelty of the paper with respect to prior work.}

	\textbf{Authors:} We addressed all the reviewers' points as discussed in the following of this document. We also emphasized the contribution of the paper with respect to our previous works in "Related work" subsection of the introduction.

@@

	\item \textbf{Associate Editor:} \textit{An updated and complete literature review should be conducted.}

	\textbf{Authors:} We performed a deep literature review. In particular, in Table 1 we classify the related work and then emphasize the novelty of our research.

@@

	\item \textbf{Associate Editor:} \textit{The relevance to Applied Energy should be enhanced with the considerations of scope and readership of the Journal.  The present form has little relevance to APEN which might suit better for another journal. Authors must address and enhance the relevance to Applied Energy otherwise, they can consider to publish in another journal.}

	\textbf{Authors:} Following the Associate Editor and reviewers comments, we re-wrote the title and the introduction of the paper focusing on the relevance of the paper to APEN. \textbf{FOR RAHUL: should we say here that this paper was invited specifically for our methodology?}

@@

	\item \textbf{Associate Editor:} \textit{A proof reading by a native English speaker should be conducted to improve both language and organization quality. }

	\textbf{Authors:} We deeply revised the paper to improve both language and organization quality.

	\item \textbf{Associate Editor:} \textit{The originality of the paper needs to be further clarified. The present form does not have sufficient results to justify the novelty of a high quality journal paper. The results should be further elaborated to show how they could be used for the real applications.}

	\textbf{Authors:} In the "Related work" and "Main contribution" subsections of the Introduction, we clarified the originality of the paper to existing results and emphasized the contributions provided. In Sections 2 and 6 we showed how our results can be used for real applications. More precisely, through a practical case study (in Section 2), we highlighted the typical drawbacks of the model-based approaches, which limit the use of advanced control optimization technique for energy optimization, such as MPC, and showed (in Sections 2 and 6) how these limitation are not a problem in the data-driven MPC approach that we propose. In particular, in Section 6, we used experimental data from a real house and considered noisy weather forecasts, to show that our approach is robust with respect to all sort of imperfections typical of the real world, i.e. error in weather predictions, unpredictable house schedules (due to people, appliancies, etc.), noisy sensor measurements (due to real experimental data), etc.. This shows that our approach can be used for real implementation.

@@

\item \textbf{Reviewer 1:} \textit{This paper presents a new framework called data predictive control which integrates machine learning and model predictive control to derive the optimal online control decisions. The framework is then applied to study several demand response problems in energy area as case studies. The problem considered in this paper is important, and the proposed approach is a good initial attempt to address the problem relying on heuristics.}

\textbf{Authors:} We thank the reviewer for the comments on the problem and approach of our paper.

\item \textbf{Reviewer 1:} \textit{In Section 2.1, what is the relationship between these training features and labels $(X^c, X^d, Y)$ and the input, output, disturbance, and state variables in the traditional state space model? Are they the same?}

\textbf{Authors:} We thank the reviewer for this comment, indeed this very important part of the paper was not clear.
We added a section (Section 2) where we reported a traditional state-space modeling of a building (Section 2.1), and compared with the data-driven modeling approach (Section 2.2). We highlighted the differences in the modeling techniques, and emphasized that the features needed in the data-driven approaches are all the measurable variables (hence a subset of all the variables) used in the model-based approach, since the model learning procedure is able to compensate the internal effect of the non-measurable ones. This is also further clarified in Section 3, where we added more mathematical details.

@@

\item \textbf{Reviewer 1:} \textit{Another point is that in traditional state space model, one cannot measure everything except input and output in most cases. How could we obtain such training data if these data are not measurable.}

\textbf{Authors:} We thank the reviewer for this comment, indeed our previous presentation of the paper was unclear on this fundamental aspect. In our revision we widely illustrate in Section 2, how in traditional state-space models part of the state is measurable (i.e. room temperatures, etc) and part is not measurable (i.e. layers/windows temperatures, etc). In our approach, of course, $(\mathcal{X}^c, \mathcal{X}^d, \mathcal{Y})$ are ALL measurable variables, i.e. variables that are present in the historical data of a building. The unmeasurable variables of the traditional state-space models are not assumed to be measurable in this paper and are neglegted in the proposed models, yet our prediction works well since we compensate them via the regression tree/random forest approaches.

\item \textbf{Reviewer 1:} \textit{The key intuition of the proposed framework is Eq. (1). It would be much better if the authors can explain their framework using Eq. (1) directly first instead of using another set of notations such as $(X^c, X^d, Y)$. The relationship between Eq. (1) and Fig. (1) is not very clear. Specifically, one would expect to estimate $g$ and $h$ using data, but how they can be trained in a two-step process as shown in Fig. 1 is not clear.}

\textbf{Authors:} We agree with the reviewer. Following this comment, and also other comments, we re-wrote the corresponding Section to make it more clear and complete to describe the DPC technical approach. In particular, we removed Eq. 1 (of the old paper), since it was confusing, and explain our goal in the beginning of Section 3. Then, in Section 3.1, we introduced the two-steps process that modifies the Regression Tree learning algorithm, to obtain models that are suitable for MPC. We clarified the relationship between the two-steps process and Fig. 1 (left), and we added a remark that explain the meaning of Fig. 1 (right).

@@

\item \textbf{Reviewer 2:} \textit{Also, what is new in section 5 compared with section 4? Even though a real house is claimed to be used, it's still tested using simulations and energyplus, like in the previous sections.}

\textbf{Authors:} We thank the reviewer for this comment, indeed the contribution of this section was not well emphasized. We improved this aspect at the beginning of the section. In particular, Section 6 provides the following differences with respect to Section 5:
\begin{itemize}
	\item In Section 5 we use the EnergyPlus model both to create the Random Forests model and to test the DPC performance.
	However, although the EnergyPlus model is quite accurate, it still provides simulated data, that differently from experimental data do not suffer of non-predictable imperfections such as random occupancy schedules, open/close windows, random light on/off switch, etc.. In section 6, instead, experimental data are used to build the Random Forests models, while the EnergyPlus model (created ad-hoc for the house) is ONLY used to make a fair comparison. Otherwise, using Random Forests models to compare DPC (that would have used the same model) and the bang-bang control would have been unfair;
	\item The optimization problems are different: in Section 5, we cut the peak of the power in a limited amount of time and require the temperature to be close to a desired value, while in Section 6 we minimize the energy usage while keeping temperature in a comfort range. Furthermore, in Section 6 we consider an on/off input, i.e. integer input variable, instead of a continuous one, showing that DPC can adapt to different problems;
	\item We introduce new simulations accounting for weather forecast uncertainties, other than the ones introduced with the experimental data, that show the robustness of DPC against disturbance errors.
\end{itemize}

\item \textbf{Reviewer 2:} \textit{A real world experiment on this house would be much more interesting.}

\textbf{Authors:} We agree with the reviewer that a real world experiment would be interesting.
However, it would not allow to compare different approaches, since they could not be ran simultaneously, hence making the comparison unfair.
We are currently working to apply our methodology on a real university building, but this is out of the scope of this paper.

### 2018-01-22 paper/final/Response_to_reviewers/Response_to_reviewers.tex

\documentclass{article}
\textwidth 6in%
\textheight 7.9in%
\oddsidemargin 1cm

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}%
\usepackage[dvips]{color}
\setcounter{MaxMatrixCols}{30}
\usepackage{natbib}

\begin{document}

\title{Authors' Response \\
\begin{Large}
	Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control
\end{Large}
}

@@

\textbf{Authors:} We thank the reviewer for rising this important issue. Yes, we assume that at time $k$ we know the future non-manipulative states (namely external temperature, radiation, etc.) for the Model Predictive Control time horizon. In the revised version of this paper, to answer the reviewer comment, we account for disturbance uncertainties and show that, even with a large error in predicting the future states, the performance does not change significantly. More precisely, we:
	\begin{itemize}
		\item Cite a study that shows, despite considering uncertainties on weather forecast on a long horizon of 72 hours, that MPC outperforms the rule-based controllers in providing energy savings, and is in general quite close to the perfect forecast case. This is due mainly to 2 main features:
		\begin{itemize}
			\item weather forecast are in general quite accurate on a shot-term horizon (especially on the length of horizon that we usually consider for DPC, for example 6 hours in Section 4, 7 hours in Section 5 and 40 minutes in Section 6);
			\item MPC strategy computes an optimal input sequence over an horizon of finite length, and then apply only the first input. At the next step, with the new EXACT measurements of the system state and of the disturbance conditions, the algorithm is applied again. This is how MPC uses feedback law to correct modeling errors;
	\end{itemize}
		\item Provide a new case study in Section 6, where we inject a noise into the disturbance and show the robustness of our approach.
	\end{itemize}

\item \textbf{Reviewer 2:} \textit{In Fig. 6b, what's the difference between blue and yellow, as they seem to completely overlap?}

\textbf{Authors:} We have clarified this point in the text and the Figure caption. Shown in yellow is the tracking signal, in blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building). The difference in closed-loop simulation and prediction is due to model mismatch.

### 2018-01-22 paper/final/abstract.tex
prefinal edits

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC) due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. 
On the other hand, data-driven models are not suitable for control because of non-physical models.
%Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. 
We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. Using a bilinear model, we show that DPC has comparable performance with MPC. We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building building model, and show that DPC curtails the desired power usage with high confidence. Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants. Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.

>>>>

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. 
Among these, one of the most popular is Model Predictive Control (MPC) due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup.
A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. 
\textcolor[rgb]{0,0,1}{To overcome this problem, we introduce a novel idea for predictive control based on historical building data leveraging machine learning algorithms like regression trees and random forests. We call this approach Data-driven model Predictive Control (DPC).
Using a physics-based bilinear model of a real building, we show that DPC provides comparable performance with MPC.
We further apply DPC to a large scale multi-story EnergyPlus building model in the problem of Demand Response, and show that DPC curtails the desired power usage with high confidence.}
Further, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for the occupants.
Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.
\textcolor[rgb]{0,0,1}{Finally, we show the robustness of DPC under disturbance uncertainties.}

### 2018-01-22 paper/final/casestudyEPlus.tex
prefinal edits

\begin{figure}[h!]
	\begin{center}
	\subfigure[Optimal inputs calculated by DPC-En. At first, the inputs are changed rapidly because of a significant difference between the desired and the actual power consumption. Then gradual adjustments are made to follow the desired reference.]{
		\label{F:eplus_inputs}
		\centering
		\includegraphics[width=29pc]{figures/eplus_control.eps}
	}
	\subfigure[Power tracking by DPC-En at 1.1 MW: The difference in closed-loop simulation and prediction is due to model mismatch.]{
		\label{F:eplus_power}
		\centering
		\includegraphics[width=29pc]{figures/eplus_closedloop.eps}
	}
\end{center}
	\caption{Power management using DPC. The controller is active between 7am - 2pm. This region is marked in dashed red lines.}
	\captionsetup{justification=centering}
	\label{F:eplus_track}
\end{figure}

@@

The results are shown in Figure~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Figure~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%. The difference between the predicted power consumption and that in the closed-loop simulation in Figure~\ref{F:eplus_power} is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}. Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

>>>>

\begin{figure}[h!]
	\begin{center}
	\subfigure[Optimal inputs calculated by DPC-En. At first, the inputs are changed rapidly because of a significant difference between the desired and the actual power consumption. Then gradual adjustments are made to follow the desired reference.]{
		\label{F:eplus_inputs}
		\centering
		\includegraphics[width=29pc]{figures/eplus_control.eps}
	}
	\subfigure[Power tracking by DPC-En at 1.1 MW. In blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building). The difference in closed-loop simulation and prediction is due to model mismatch.]{
		\label{F:eplus_power}
		\centering
		\includegraphics[width=29pc]{figures/eplus_closedloop.eps}
	}
\end{center}
	\caption{Power management using DPC. The controller is active between 7am - 2pm. This region is marked in dashed red lines.}
	\captionsetup{justification=centering}
	\label{F:eplus_track}
\end{figure}

@@

The results are shown in Figure~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Figure~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%.
\textcolor[rgb]{0,0,1}{The results from the closed-loop simulation are shown in Figure~\ref{F:eplus_power}.
Marked in blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
The difference between the predicted power consumption and that in the closed-loop simulation is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.} 
Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

### 2018-01-22 paper/final/casestudyReal.tex
prefinal edits

In this section, we show an application of DPC to a real house located in L'Aquila, Italy. Random forest models for DPC built using historical data from this house are described in Section \ref{SS:randomForestsModels}. 
In Section \ref{SS:energyPlusmodel}, we present the thermal energy model of the building in EnergyPlus, built using historical data, construction layout and materials after spending $\sim3$ months of efforts.
In Section \ref{SS:controllersDPCandBangbang}, we set up the DPC optimization problem and describe the bang-bang control strategy.
Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy saving using the EnergyPlus model.
\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC allows an energy saving up to $49.2\%$ with respect to the bang-bang controller while guaranteeing thermal comfort for occupants, considering the perfect knowledge of the weather forecast. 
In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
The section organization is graphically shown in Figure \ref{F:overview}.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller. DPC optimzation uses Random Forests with affine functions as predictive models. Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}

\textcolor[rgb]{0,0,1}{The contribution of this section with respect to Sections \ref{S:casestudy} is threefold:

@@

	\item we consider experimental data from a lived house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc..
	On the other hand, the EnergyPlus model considered in Section \ref{S:casestudy}, although provides an accurate model of a building, is still a simulated model that does not take into account the aforementioned imperfections;
	\item we modify the DPC algorithm to take into account uncertainties on the weather forecast, so making DPC robust with respect to prediction errors;
	\item in Section \ref{S:casestudy}, we want to cut the peak of the power in a specific time range while keeping the temperature as close as possible to a specific setpoint.
	Differently, in this section, we show the adaptability of DPC to different problems, minimizing the energy consumption required to keep room temperatures within a specified range of comfort for an arbitrary amount of time. In this way we show the potential of DPC for energy saving on a long period. 

@@

We ran DPC with $3$ different sets of parameters $Q$, $\lambda_{min}$ and $\lambda_{max}$. Each set allows a different level of temperature bounds violation. In particular, we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
\textcolor[rgb]{0,0,1}{The simulation period is of $15$ days, from May $1$, $2016$ at $00am$ to May $15$, $2016$ $00am$.}

\textcolor[rgb]{0,0,1}{We considered 2 different simulative conditions to show the robustness of our approach:
\begin{enumerate}
	\item in Section \ref{SSS:DisturbancePerfect} we provide our simulative results considering perfect knowledge of the weather forecast;
	\item in Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainty Hence we modify Problem \ref{P:dpcRealCase} by adding a gaussian noise with high variance on the perfect forecast, and show that the results are close to the perfect forecast case, so making our methodology robust with respect to disturbance uncertainties.
\end{enumerate}
\subsubsection{Perfect knowledge of the weather forecast}\label{SSS:DisturbancePerfect}
In this section we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
\paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}. For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from May $1$, $2016$ $00am$ to May $5$, $2016$ $1pm$. \textcolor[rgb]{0,0,1}{The whole period will be used in "Result 2" to provide the bounds violation errors, and in "Result 3" for the energy consumption comparison.}
We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$. Bang-bang control also presents small bounds violations due to its working principle. We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range. 
We will see in Figure \ref{F:comparison_all_energy_E+} that this translates to significant energy saving.

>>>>

In this section, we show an application of DPC to a real house located in L'Aquila, Italy. Random forest models for DPC built using historical data from this house are described in Section \ref{SS:randomForestsModels}. 
In Section \ref{SS:energyPlusmodel}, we present the thermal energy model of the building in EnergyPlus, built using historical data, construction layout and materials after spending $\sim3$ months of efforts.
In Section \ref{SS:controllersDPCandBangbang}, we set up the DPC optimization problem and describe the bang-bang control strategy.
Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy savings using the EnergyPlus model.
\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC provides up to $49.2\%$ energy savings with respect to the bang-bang controller while guaranteeing thermal comfort for the occupants, considering the perfect knowledge of the weather forecast. 
In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
The section organization is graphically shown in Figure \ref{F:overview}.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller. DPC optimzation uses Random Forests with affine functions as predictive models. Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}

\textcolor[rgb]{0,0,1}{The contribution of this section is threefold.

@@

	\item We consider experimental data from an occupied house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc.
	This is different from Section \ref{S:casestudy} where we used data simulated from an EnergyPlus model.
	\item We modify the DPC algorithm to take into account uncertainties In the weather forecast, making DPC robust with respect to prediction errors.
	\item While in Section \ref{S:casestudy} we follow a reference DR signal while maintaining thermal comfort, in this section we show the adaptability of DPC to different problems where we minimize the energy consumption required to keep the room temperatures within a specified range of comfort. In this way, we show the potential of DPC in providing energy savings for a long period. 

@@

We ran DPC with $3$ different sets of parameters $Q$, $\lambda_{min}$ and $\lambda_{max}$. Each set allows a different level of temperature bounds violation. In particular, we considered a small ($Q=100$, $\lambda_{min}=3000$ and $\lambda_{max}=100$), a medium ($Q=100$, $\lambda_{min}=1000$ and $\lambda_{max}=100$) and a large ($Q=100$, $\lambda_{min}=100$ and $\lambda_{max}=100$) violation configuration.
\textcolor[rgb]{0,0,1}{The simulation period is of $15$ days, from $00$ hrs on May $1$, $2016$ to $00$ hrs on May $15$, $2016$.}

\textcolor[rgb]{0,0,1}{We considered two different simulation conditions to show the robustness of our approach.
\begin{enumerate}
	\item In Section \ref{SSS:DisturbancePerfect}, we provide the results considering perfect knowledge of the weather forecast.
	\item In Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainties. Hence, we modify Problem \ref{P:dpcRealCase} by adding a Gaussian noise with high variance to the perfect forecast, and show that the results are close to the perfect forecast case.
	This verifies our methodology is robust with respect to disturbance uncertainties.
\end{enumerate}
\subsubsection{Perfect knowledge of the weather forecast}\label{SSS:DisturbancePerfect}
In this section, we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
\paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}. For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from $00$ hrs on May $1$, $2016$ to  $13$ hrs on May $5$, $2016$. \textcolor[rgb]{0,0,1}{The whole period will be used in ``Result 2" to provide the bounds violation errors, and in ``Result 3" for the energy consumption comparison.}
We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$. Bang-bang control also presents small bounds violations due to its working principle. We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range. 
We will see in Figure \ref{F:comparison_all_energy_E+} that this translates to significant energy saving.

### 2018-01-22 paper/final/conclusion.tex
prefinal edits

To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data-driven model Predictive Control (DPC). Data-based control strategies are usually based on non-physical models, therefore the can not be integrated with most of the classical control approaches. The goal is to create data-driven models that are suitable for receiding horizon control. To this aim, we present two algorithms, based on trees and random forests, to create data-driven models to be used in the DPC strategy. We then apply DPC to $3$ different case studies in order to prove its strength.
\begin{enumerate}
\item \emph{\textbf{Comparison with MPC.}} We compare the performance of our DPC to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
\item \emph{\textbf{Demand Response application.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
\item \emph{\textbf{\textcolor[rgb]{0,0,1}{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses. We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants. With respect to a bang-bang controller, we obtained an energy saving that goes from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations of such range. Furthermore, we also demonstrate the robustness of DPC considering uncertainties on weather forecast.}
\end{enumerate}

>>>>

To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data-driven model Predictive Control (DPC). \textcolor[rgb]{0,0,1}{Data-driven control is based on non-physical (black-box) models, therefore they can not be integrated with most of the classical control approaches.} The goal is to create data-driven models that are suitable for receiding horizon control. \textcolor[rgb]{0,0,1}{To this aim, we present two algorithms, based on trees and random forests, to create control-oriented models for DPC. We then apply DPC to three different case studies to demonstrate its strength.}
\begin{enumerate}
\item \emph{\textbf{Comparison with MPC.}} We compare the performance of our DPC to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
\item \emph{\textbf{Demand Response application.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
\item \textcolor[rgb]{0,0,1}{\emph{\textbf{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control. We show that DPC provides energy savings while guaranteeing thermal comfort for the occupants. With respect to a bang-bang controller, we obtain energy savings that go from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations in the thermal comfort level. Furthermore, we demonstrate the robustness of DPC considering uncertainties in the weather forecast.}
\end{enumerate}

### 2018-01-22 paper/final/cs-challenges.tex
prefinal edits

\begin{enumerate}
\item \emph{\textbf{Data Availability:}} The main practical challenge for DPC lies in the availability of data for training, and we require answers to questions like \emph{how much data (functional testing) is required, and how should the sampling be done?} Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work.
\item \emph{\textbf{Stability:}} While the buildings are inherently stable, many other applications require stability guarantees. In our ongoing work, we are working towards proving asymptotic stability to origin with DPC-RT and DPC-En by using concept of switched LTI systems. This will make DPC useful for systems with faster dynamics.
\end{enumerate}

>>>>

\begin{enumerate}
\item \emph{\textbf{Data Availability:}} \textcolor[rgb]{0,0,1}{The main practical challenge for DPC lies in the availability of data for training, and we require answers to questions like \emph{how much data is required, and how should the sampling (functional testing) be done?} Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work \cite{JainICCPS2018}}.
\item \emph{\textbf{Stability:}} While the buildings are inherently stable, many other applications require stability guarantees. In our ongoing work, we are working towards proving asymptotic stability to origin with DPC-RT and DPC-En by using concept of switched LTI systems. This will make DPC useful for systems with faster dynamics.
\end{enumerate}
